---
title: "230724_TCGAvsUVM"
author: "Haroon Riyaz"
date: "2023-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```




~~~~~~~~~
#Packages
~~~~~~~~~

```{r}
library(readxl)
library(data.table)
```




~~~~~~~~~~~~~~~~
#TCGA Tumor Data
~~~~~~~~~~~~~~~~


##Formatting

Load Log2 gene expression set (normalization already done through log2(count+1))
```{r}
XENAexpression <- readr::read_tsv("TCGA-UVM.htseq_counts.tsv") 
```

Format expression file
```{r}
XENAexpression <- as.data.frame(XENAexpression) 

# Removes "-01A" text from all column names 
XENAcolnames <- colnames(XENAexpression)
formatted_xena_colnames <- sub("-01A", "", XENAcolnames)
formatted_xena_colnames <- sub("-01B", "", formatted_xena_colnames)
formatted_xena_colnames <- sub("-", ".", formatted_xena_colnames)
formatted_xena_colnames <- sub("-", ".", formatted_xena_colnames)

colnames(XENAexpression) <- formatted_xena_colnames
```

Makes gene names the row-names
```{r}
rownames(XENAexpression) <- XENAexpression[,1]
XENAexpression <- XENAexpression[,-1]
```

Format annotation data 
```{r}
XENAannotation <- as.data.frame(read_xlsx("TCGA_BAP1.xlsx"))
rownames(XENAannotation) <- XENAannotation[,1]
```


##Without DST_MTHD


###Slingshot 

```{r}
glab_Slingshot(expression = XENAexpression, meta = XENAannotation, meta_trait = "Group No.", start.clus = "1", nfeatures = 5000, npcs = 10)
```


###Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 2, PCX2 = 3, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 3, text_size = 1.5)
```


###3D Plot

```{r}
glab_Slingshot_plot_3D(sds = sds, meta = meta, meta_trait = meta_trait)
```


##Scaled Distance Method


###Slingshot 

```{r}
glab_Slingshot(expression = XENAexpression, meta = XENAannotation, meta_trait = "Group No.", start.clus = "1", nfeatures = 5000, npcs = 10, dist.method = "scaled.diag")
```


###Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 2, PCX2 = 3, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 3, text_size = 1.5)
```

###3D Plot

```{r}
glab_Slingshot_plot_3D(sds = sds, meta = meta, meta_trait = meta_trait)
```




~~~~~~~~~~~~~~~~~~~~
#SKCM Cell-line Data
~~~~~~~~~~~~~~~~~~~~


##Formatting 

Load Log2 gene expression set (normalization already done through log2(count+1))
```{r}
SKCMexpression <- data.frame(fread("SKCM_CellLines.txt"))
```

Format expression file
```{r}
rownames(SKCMexpression) <- SKCMexpression$gene   
  SKCMexpression <- SKCMexpression[,-1] 
```

Format annotation data 
```{r}
SKCMannotation <- read.table("SKCM_group_annotation.txt")  
SKCMannotation[20:21,2] <- NA # Make UPMD1 and WM3772F NA
  
rownames(SKCMannotation) <- SKCMannotation[,1]
```


##Without DST_MTHD


###Slingshot 

```{r}
glab_Slingshot(expression = SKCMexpression, meta = SKCMannotation, meta_trait = "Group", start.clus = "SKCM4", nfeatures = 5000, npcs = 10)
```


###Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 2, PCX2 = 3, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 3, text_size = 1.5)
```


###3D Plot

```{r}
glab_Slingshot_plot_3D(sds = sds, meta = meta, meta_trait = meta_trait)
```


##Scaled Distance Method


###Slingshot 

```{r}
glab_Slingshot(expression = SKCMexpression, meta = SKCMannotation, meta_trait = "Group", start.clus = "SKCM4", nfeatures = 5000, npcs = 10, dist.method = "scaled.diag")
```


###Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 2, PCX2 = 3, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 3, text_size = 1.5)
```


###3D Plot

```{r}
glab_Slingshot_plot_3D(sds = sds, meta = meta, meta_trait = meta_trait)
```




~~~~~~~~~~~~~~~~~~~
#UVM Cell-line Data
~~~~~~~~~~~~~~~~~~~


##Formatting 

Load Log2 gene expression set (normalization already done through log2(count+1))
```{r}
UVMexpression <- fread("UVM_UCLA_BROAD_geneexp_22Q4_pc_genes_log2.csv") 
```

Format expression file
```{r}
UVMexpression <- as.data.frame(UVMexpression) 
```

Makes gene names the row-names
```{r}
UVMexpression <- UVMexpression[-which(duplicated(UVMexpression$gene)),]
rownames(UVMexpression) <- UVMexpression$gene
UVMexpression <- UVMexpression[,-1]
```

Formats annotation data 
```{r}
UVMannotation <- read.csv("UVM_UCLA_BROAD_22Q4_annotation_file.csv")
rownames(UVMannotation) <- UVMannotation[,1]
UVMannotation$cellline <- 0
```

* OTHER FORMATTING * Different grouping for UVM cell-lines
```{r}
newUVMannotation <- UVMannotation
newUVMannotation[c("MEL202.BROAD", "MEL202.UCLA", "X921.BROAD", "X92.1.UCLA"),"Group"] <- 1
newUVMannotation[c("MP41.UCLA", "OMM1.BROAD"),"Group"] <- 2
newUVMannotation[c("UPMD1.BROAD", "MEL270.BROAD"),"Group"] <- 3
newUVMannotation[c("MEL070.UCLA", "MEL270.repeat.UCLA", "OMM1.5.UCLA", "OMM1.3.UCLA", "OMM25.BROAD"),"Group"] <- 4
newUVMannotation[c("MEL285.BROAD", "MEL290.BROAD"),"Group"] <- 5
newUVMannotation[c("MP46.BROAD", "WM3772F.BROAD"),"Group"] <- 6
newUVMannotation[c("MM28.UCLA", "MEL196.UCLA"),"Group"] <- 7
newUVMannotation[c("MEL045.UCLA", "MP38.UCLA"),"Group"] <- 8
```


##Without DST_MTHD


###Slingshot 

The method for calculating distances between clusters normally used by Slingshot does not work with 10 PCs. Using "scaled.diag" instead fixes this issue. 

```{r}
glab_Slingshot(expression = UVMexpression, meta = UVMannotation, meta_trait = "Group", start.clus = "4c", nfeatures = 5000, npcs = 10)
```


##Scaled Distance Method


###Slingshot 

```{r}
glab_Slingshot(expression = UVMexpression, meta = UVMannotation, meta_trait = "Group", start.clus = "4c", nfeatures = 5000, npcs = 10, dist.method = "scaled.diag")
```


###Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 2, PCX2 = 3, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 3, text_size = 1.5)
```


###3D Plot

```{r}
glab_Slingshot_plot_3D(sds = sds, meta = meta, meta_trait = meta_trait)
```


## OTHER FORMATTING 


###Slingshot

```{r}
glab_Slingshot(expression = UVMexpression, meta = newUVMannotation, meta_trait = "Group", start.clus = "1", nfeatures = 5000, npcs = 10, dist.method = "scaled.diag")
```


###Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 2, PCX2 = 3, text_size = 1.5)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 3, text_size = 1.5)
```


###3D Plot

```{r}
glab_Slingshot_plot_3D(sds = sds, meta = meta, meta_trait = meta_trait)
```





