---
title: "230724_TCGAvsUVM"
author: "Haroon Riyaz"
date: "2023-07-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages

```{r}
library(readxl)
library(data.table)
```


#TCGA Tumor Data

Load Log2 gene expression set (normalization already done through log2(count+1))

```{r}
XENAexpression <- readr::read_tsv("TCGA-UVM.htseq_counts.tsv") 
```

Format expression file

```{r}
XENAexpression <- as.data.frame(XENAexpression) 

# Removes "-01A" text from all column names 
XENAcolnames <- colnames(XENAexpression)
formatted_xena_colnames <- sub("-01A", "", XENAcolnames)
formatted_xena_colnames <- sub("-01B", "", formatted_xena_colnames)
formatted_xena_colnames <- sub("-", ".", formatted_xena_colnames)
formatted_xena_colnames <- sub("-", ".", formatted_xena_colnames)

colnames(XENAexpression) <- formatted_xena_colnames
```

Makes gene names the row-names

```{r}
rownames(XENAexpression) <- XENAexpression[,1]
XENAexpression <- XENAexpression[,-1]
```

Formats annotation data 

```{r}
XENAannotation <- as.data.frame(read_xlsx("TCGA_BAP1.xlsx"))
rownames(XENAannotation) <- XENAannotation[,1]
```

##Slingshot 

```{r}
glab_Slingshot(expression = XENAexpression, meta = XENAannotation, meta_trait = "Group No.", start.clus = "1", nfeatures = 5000, npcs = 10)
```

##Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```


*
#UVM Cell-line Data
*

Load Log2 gene expression set (normalization already done through log2(count+1))

```{r}
UVMexpression <- fread("UVM_UCLA_BROAD_geneexp_22Q4_pc_genes_log2.csv") 
```

Format expression file

```{r}
UVMexpression <- as.data.frame(UVMexpression) 
```

Makes gene names the row-names

```{r}
UVMexpression <- UVMexpression[-which(duplicated(UVMexpression$gene)),]
rownames(UVMexpression) <- UVMexpression$gene
UVMexpression <- UVMexpression[,-1]
```

Formats annotation data 

```{r}
UVMannotation <- read.csv("UVM_UCLA_BROAD_22Q4_annotation_file.csv")
rownames(UVMannotation) <- UVMannotation[,1]
UVMannotation$cellline <- 0
```

* OTHER FORMATTING 

```{r}
new_meta <- UVMannotation
new_meta[c("MEL202.BROAD", "MEL202.UCLA", "X921.BROAD", "X92.1.UCLA"),"Group"] <- 1
new_meta[c("MP41.UCLA", "OMM1.BROAD"),"Group"] <- 2
new_meta[c("UPMD1.BROAD", "MEL270.BROAD"),"Group"] <- 3
new_meta[c("MEL070.UCLA", "MEL270.repeat.UCLA", "OMM1.5.UCLA", "OMM1.3.UCLA", "OMM25.BROAD"),"Group"] <- 4
new_meta[c("MEL285.BROAD", "MEL290.BROAD"),"Group"] <- 5
new_meta[c("MP46.BROAD", "WM3772F.BROAD"),"Group"] <- 6
new_meta[c("MM28.UCLA", "MEL196.UCLA"),"Group"] <- 7
new_meta[c("MEL045.UCLA", "MP38.UCLA"),"Group"] <- 8
```

##Slingshot 

```{r}
glab_Slingshot(expression = UVMexpression, meta = UVMannotation, meta_trait = "Group", start.clus = "4c", nfeatures = 5000, npcs = 2)
```
##Plot

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```

* OTHER FORMATTING 

```{r}
glab_Slingshot(expression = UVMexpression, meta = new_meta, meta_trait = "Group", start.clus = "4c", nfeatures = 5000, npcs = 2)
```

```{r}
glab_Slingshot_plot(rd = rd, clusters = clusters, curves = curves, MST = MST, PCX1 = 1, PCX2 = 2, text_size = 1.5)
```






